Can government public communications elicit undue trust? Exploring the interaction between symbols and substantive information in communications

### 2019-03-15 paper2_JPART_rev_01.Rmd
Add files via upload

Public administration scholars are highly concerned about the implications of citizens’ low trust in government, and their underestimation of public-service performance. ##ADD more about trust.##

@@

Extant research on the effect of government public communications on citizens attitudes has mostly focused on the *informative* aspects of these communications, whereby government organizations provide citizens with evidence about their performance and details and explanations about their policies and actions [e.g. @marvel_2015; @im_2014; @porumbescu_2017; @porumbescu_2016; @porumbescu_2018; @grimmelikhuijsen_2015; @ho_2017]. This theoretical and empirical focus on information and transparency congrues with the normative view of government public communication as a means for accomplishing democratic accountability, facilitating the effective implementation of their missions, and aligning citizens' perceptions of public organizations with their actual performance [@canel_2013; @lee_2012; @lee_2001; @lee_2008; @lee_2011]. According to this view, public communications are expected to increase citizens' trust in bureaucracies mainly insofar as these communications provide citizens with favorable information on the performance of organizations and logically persuasive explanations for their policies.

Recently, a few studies pointed to the additional effect of *symbolic* elements that are entangled in government communications, such as  agencies' names and brand logos, figures, images and celebrity endorsements [@alonbarkat_2017; @karens_2016; @teodoro_2018]. These studies demonstrated that government organizations may cause citizens to view them more favorably through familiar and well-designed symbolic elements that evoke positive associations and emotions. Still, we lack a theory and robust empirical evidence of the exact mechanisms through which these symbols increase citizens’ trust. These previous studies also did not examine the interaction between the symbolic elements and the substantive content of information in public communications.

This research gap has normative significance. Whereas the responsiveness of citizens to substantive information is perceived favorably from a democratic theory standpoint, the effect of symbolic elements can be problematic. One can argue that shaping citizens' perceptions of governments by evoking associations and emotions stands against the normative expectation that citizens' judgments should be based on their critical thinking. We need to be aware that symbolic communication may elicit undue trust in poorly performing organizations and in ill-conceived policies, and enable public organizations to escape justified public criticism [@alonbarkat_2017]. Therefore, this study focuses on the following questions: *What are the psychological mechanisms by which symbolic elements in government communications affect citizens' trust in government organizations and their policies? Does the effect of symbolic elements depend on the substantive content of communications?* *Do symbols offset the rational, negative effect of unfavorable or logically unpersuasive information?*

@@

To empirically test these theoretical expectations, I conducted a large randomized survey experiment among circa `r SVIVA2_01%>%nrow()%>%round(-1)` Israeli citizens. The experiment focuses on the empirical case of the Israeli Environment Protection Ministry. I explore citizens' responses to public communications about two policy plans of the Ministry, and specifically the extent to which their trust in these policies are affected by the substantive information in the communications (logically persuasive vs. unpersuasive policy plans) and by symbolic elements in them (real symbols vs. fake symbols and no symbols conditions).

I found that the symbolic elements increased citizens' trust in these policies. In line with my theoretical expectations, the effect was greater among those who saw the logically unpersuasive policy plans. They were also less inclined to think about the policy content because they were distracted by the symbols. I discuss the theoretical and normative implications of these findings.

@@

Symbols are fundamental for almost any form of human interaction, as acknowledged by early scholars of social science [@blumer_1986]. A symbol can be broadly defined as *"any object used by human beings to index meanings that are not inherent in, nor discernible from, the object itself"* [@elder_1983, pp. 28-29]. My focus is on those symbols that are strategically used by government organizations in their public communications, are designed to distinguish them and to evoke associations and emotions that stimulate audiences’ positive affect towards organizations and their operations. Prominent examples are unique logos, colors associated with specific meanings, icons, images, and the use of celebrities as endorsers. These elements are widely used by businesses and political candidates to attract targeted consumers and voters. Yet, they can also be found in the communications of government bureaucracies with citizens, including their mass-media advertising campaigns, social media accounts, internet pages, official publications, signboards in public places and personal mail notifications to citizens.

In comparison with the abundant research on citizens’ responses to informative elements in public communications [for reviews see: @cucciniello_2016; @grimmelikhuijsen_2017; @james_2017], public administration literature has devoted relatively little attention to the effect of symbolic elements. Recently, however, a few experimental studies tackled the issue. @marvel_2015 analyzed US citizens' responses to a television advertisement of the public US Postal Service, which involved various symbolic elements, including the organization's logo, workers' images and background music. His study shows that participants who watched the advertisement evaluated the organization's performance more positively. @teodoro_2018 suggested that the names of US federal agencies not only represent information about their activities, but may also carry specific symbolic meaning and positive/negative "affect" that can shape citizens' support for government policies (depending on citizens' political party identification). While these two studies provide examples for the role of symbols in shaping citizens' attitudes, they did not effectively disentangle the effects of the symbolic and informational aspects of the public communications studied.

A study by @karens_2016 did examine the distinct effect of symbols in communications on citizens' trust. The scholars surveyed university students from three European countries and examined their trust in two policy plans of the European Commission, and it's shaping by the appearance of the organization's logo in the communication of the policy plans. They found that including the Commission’s logo significantly increased citizens' trust in these policies. Their results show an impressive effect size equivalent to 0.7 standard deviations, on average.

@@

Even though these studies substantiate a link between the appearance of symbolic elements in communications and citizens' favorable attitudes, they did not directly investigate the causal mechanisms that underlie this effect. Therefore, we still lack understanding of the circumstances under which symbols are likely to be more/less effective in shaping citizens' trust. Specifically, these studies neither theorized nor empirically tested the interaction between symbolic elements and citizens’ processing of substantive information in communications. We do not know whether well-designed symbols reinforce, or offset, citizens' rational responses to information. Specifically, as explained above, it is essential to assess the possibility that symbols can unduly compensate for unfavorable or logically unpersuasive information that should yield negative responses.

@@

The social psychology Elaboration Likelihood Model (ELM) provides a comprehensive theoretical framework for investigating peoples' varying responses to "persuasive" communications [@petty_1986; @petty_2011]. In addition to its general theoretical application in psychological research, this model has been applied to studies on persuasion in various areas, including commercial advertising [e.g. @petty_1983; @dotson_2000; @wheeler_2005; @moore_1987], public health campaigns [@wilson_2007; @angst_2009; @ruiter_2001; @briol_2006], media coverage [@petty_2002; @vanderwurff_2016] and media entertainment [@young_2008; @slater_2002]. So far however the relevance of ELM has not been directly examined in relation to citizens' responses to government communications.

According to ELM, the responses of individuals to persuasive communications depends on how much effort they devote to thinking about the content of the message (or "the amount of thinking" according to ELM terminology). Persuasion can occur when thinking is "high" or "low", but the processes and consequences are likely to differ in each situation. This range between high and low degrees of thought is referred to as the "elaboration continuum". People's location along this continuum is associated with different psychological processes of persuasion. At the high end of the elaboration continuum, also called the "central route", people are likely to respond by cognitively scrutinizing the arguments central to the merits of the issue. At the low end, also called "the peripheral route" people respond to simple cues automatically and unconsciously through primitive psychological processes that require very little cognitive effort, such as an unconscious transfer of affect[^footnote_dual_process_theories]. In turn, these different persuasion mechanisms are likely to cause divergent types of attitudinal change. Persuasion via the central route is expected to be more enduring, resistant and predictive of behavior, compared with the peripheral route. Of course, often people are located in between the two ends of the continuum, and thus persuasion is determined by a combination of these processes.

@@

ELM enables us to disentangle the distinct effects of different elements in communications. Some elements are processed differently by people at different locations along the elaboration continuum. Consider for instance a text that contains a series of logical arguments supporting a certain position. Those at the high end of the continuum would regard this text as argumentation, systematically scrutinizing it based on its logical persuasive quality (central route processing). Conversely, those at the low end of the continuum are more likely to disregard the content of the text, and instead, rely on the number of lines in the text, for example, as a cue for its strength (peripheral route processing). Some elements provide potential cues that lack any logical quality at all. For example, the physical attractiveness of the source normally does not represent any substantive argument, yet it can serve as a simple cue. These latter elements may affect the attitudes of those at the low end, but are less likely to persuade those at the high end [@petty_1986].

@@

How may we apply the theoretical framework of ELM to the study of the effects of symbolic elements in government communications with citizens and their interaction with substantive information?

In light of the conceptualization of ELM, substantive information elements in government public communications represent issue-related arguments, which can be scrutinized or processed systematically based on their logical quality (via the central route), at least by those who are sufficiently motivated and able to do so. Conversely, symbolic elements are potential peripheral cues, which can affect citizens' trust via the peripheral route, by evoking positive emotions and associations that are attributed to government organizations and their policies [@alonbarkat_2017]. However, in addition to their direct effect as peripheral cues, symbolic elements may also increase citizens' trust indirectly. They can decrease the amount of thought citizens dedicate to the communication, and thus prevent them from scrutinizing negative information (e.g. information on poor performance) and weak arguments (e.g. logically unpersuasive description of their policy plans).[^footnote_interaction_strong_information]

Symbolic elements may decrease citizens' ability and motivation to think about the message in several ways. First, they can cause elaboration disruption, by distracting citizens and capturing their attention and cognitive efforts [@petty_1976; @petty_1986]. Second, they can decrease citizens' motivation and ability to think about the message content by creating a positive affect and placing citizens in a positive mood. As explained above, the postulation that positive mood decreases elaboration has been explained, in part, by a mood maintenance mechanism [@mackie_1989; @bless_1990; @schwarz_1991]. This mechanism is particularly relevant to the context of government communications, in which thinking about the content of the message – government organizations and policies – is likely to yield a negative mood among most people [@marvel_2015; @marvel_2015a; @hvidman_2016]. Therefore, when citizens are exposed to symbolic elements that evoke positive feelings, they may seek to avoid thinking about tedious bureaucratic affairs, which would depress their positive mood. The latter expectation entails an interaction between symbols and the substantive information in the communications.

@@

@@

Keeping these limitations in mind, this study has several important implications for public administration research. Most importantly, my findings highlight the significant role of branding and symbolic communication in shaping citizens' attitudes about government. Citizens' responses to symbolic elements in communications thus far received relatively little attention in Public administration literature, compared with the abundant research on their responses to information. Yet, this study, and others, demonstrated that symbols can be fairly effective in increasing citizens' trust, and mitigating their negative attitudes [@alonbarkat_2017; @karens_2016; @teodoro_2018].

Furthermore, my findings suggest that symbols decrease citizens' "rational", systematic processing of substantive information. This is good news as far as we are concerned mainly about citizens' *undervaluation* of government performance, and the negative biases that undermine the effectiveness of positive information [@baekgaard_2016; @hvidman_2016; @james_2014; @james_2017; @marvel_2015; @olsen_2017; @james_2017a; @james_2017b; @christensen_2018; @hvidman_2018]. On the other hand, the emotive effect of government symbols on public opinion is highly problematic, insofar as it challenges the democratic ideal that citizens form their opinions about their government and its actions on the basis of rational thinking [@dahl_1973; @soroka_2010]. Moreover, it becomes especially problematic when symbols are used by poorly performing organizations, or organizations that lack professional expertise. In those cases, symbolic elements can mitigate justified public criticism of government failures. The potency to manipulate public opinion via symbols could be used for promoting the narrow political interests of parties and individual politicians controling the bureaucracies, as suggested by case studies of governemnt public relations [@holtzbacha_2013; @sanders_2013; @young_2013]. At the same time, it could also be used for promoting the interests of the bureaucracies themselves, namely nurturing their long-term positive reputation, which provides them with political power and autonomy [@busuioc_2016; @bertelli_2018]. Hence, whereas government communication with citizens is normally perceived as a tool for enhancing democratic responsiveness and accountability, it can also be use it as a means to undermine them.

@@

>>>>

Scholars of Political-Science and Public-Administration are long concerned about citizens’ low trust in government, and their underestimation of public-service performance [@alford_2001; @nye_1997; @levi_2000; @vandewalle_2008; @vigodagadot_2006; @vanryzin_2011]. Distrust in public organizations has a negative impact of government performance and capacity. It may hamper citizens' collaboration with organizations' policies, which, in turn, affect their successful implementation. It is also said to encourage citizens' tax evasion and selection of private over public service providers, and it could make it more difficult for public sector organizations to attract and keep high-quality employees [@hvidman_2018].

@@

Extant research on the effect of government public communications on citizens trust has mostly focused on the *informative* aspects of these communications, whereby government organizations provide citizens with evidence about their performance and details and explanations about their policies and actions [e.g. @marvel_2015; @im_2014; @porumbescu_2017; @porumbescu_2016; @porumbescu_2018; @grimmelikhuijsen_2015; @ho_2017]. This focus on information and transparency is in line with the positive normative view of government public communication and public relations, as articulated by the public administration literature on this topic, led by the works of Merdechai Lee. Communications with citizens are perceived as a means for accomplishing democratic accountability, facilitating the effective implementation of their missions, and aligning citizens' perceptions of public organizations with their actual performance [@canel_2013; @lee_2000; @lee_2012; @lee_2001; @lee_2008; @lee_2011]. Importantly, according to this view, public communications are expected to increase citizens' trust in bureaucracies mainly insofar as these communications provide citizens with favorable information on the performance of organizations and logically persuasive explanations for their policies.

Recently, however, a few studies pointed to the additional effect of *symbolic* elements that are entangled in government communications, such as  agencies' names and brand logos, figures, images and celebrity endorsements [@alonbarkat_2017; @karens_2016; @teodoro_2018]. These studies experimentally demonstrated that government organizations may cause citizens to view them more favorably and trust them through familiar and well-designed symbolic elements that evoke positive associations and emotions. Still, we lack a theory and robust empirical evidence of the exact mechanisms through which these symbols increase citizens’ trust. These studies also did not examine the interaction between the symbolic elements and the substantive content of information in public communications.

This research gap has normative significance. Whereas the responsiveness of citizens to substantive information is perceived favorably from a democratic standpoint (assuming that it is accurate), establishing trust through symbolic elements can be highly problematic. Shaping citizens' perceptions of governments by evoking associations and emotions stands against the normative expectation that citizens' judgments should be based on their critical thinking. Furthermore, we need to be aware that symbolic communication may elicit undue trust in poorly performing organizations and in ill-conceived policies, and enable public organizations to escape justified public criticism [@alonbarkat_2017]. Therefore, this study focuses on the following questions: *What are the psychological mechanisms by which symbolic elements in government communications affect citizens' trust in government organizations and their policies? Does the effect of symbolic elements on trust depend on the substantive content of communications?* *Do symbols offset the rational, negative effect of unfavorable or logically unpersuasive information?*

@@

To empirically test these theoretical expectations, I conducted a large randomized survey experiment among circa `r SVIVA2_01%>%nrow()%>%round(-1)` Israeli citizens. The experiment focuses on the empirical case of the Israeli Environment Protection Ministry. I explore citizens' responses to public communications about two policy plans of the Ministry. Specifically, I examine how their trust in these public policies is affected by the symbolic elements in the communications (real symbols vs. fake symbols and no symbols conditions) as well as by the substantive information in them (logically persuasive vs. unpersuasive policy plans).

I found that the symbolic elements increased citizens' trust in these policies. In line with my theoretical expectations, the effect was greater among those who saw the logically unpersuasive policy plans. The latter were less inclined to think about the weak, unpersuasive content, because they were distracted by the symbols. I discuss the theoretical and normative implications of these findings.

@@

A symbol can be broadly defined as *"any object used by human beings to index meanings that are not inherent in, nor discernible from, the object itself"* [@elder_1983, pp. 28-29]. Symbols are fundamental for almost any form of human interaction, and specifically in the context of politics and public affairs [@edelman_1964; @edelman_1988; @elder_1983]. Conselletions of symbols in discourse form consolidated narratives, which play an key role in shaping people's perceptions of policy problems and solutions [@miller_2012; @miller_2015].

My focus in this study is on those symbols that are strategically used by government organizations in their public communications, are designed to distinguish them and to evoke associations and emotions that stimulate audiences’ positive affect towards organizations and their operations. Prominent examples are unique logos, colors associated with specific meanings, icons, images, and the use of celebrities as endorsers. These elements are widely used by businesses and political candidates to attract targeted consumers and voters. Yet, they can also be found in the communications of government bureaucracies with citizens, including their mass-media advertising campaigns, social media accounts, internet pages, official publications, signboards in public places and personal mail notifications to citizens.

In comparison with the abundant research on the effect of informative elements in public communications on public trust [for reviews see: @cucciniello_2016; @grimmelikhuijsen_2017; @james_2017], public administration literature has devoted relatively little attention to the effect of symbolic elements. Recently, however, a few experimental studies tackled the issue. @marvel_2015a analyzed US citizens' responses to a television advertisement of the public US Postal Service, which involved various symbolic elements, including the organization's logo, workers' images and background music. His study shows that participants who watched the advertisement evaluated the organization's performance more positively. @teodoro_2018 suggested that the names of US federal agencies not only represent information about their activities, but may also carry specific symbolic meaning and positive/negative "affect" that can shape citizens' support for government policies (depending on citizens' political party identification). While these two studies provide examples for the role of symbols in shaping citizens' trust and performance evaluation, they did not effectively disentangle the effects of the symbolic and informational aspects of the public communications studied.

A study by @karens_2016 did examine the distinct effect of symbols in communications on citizens' trust in bureaucracies. The scholars surveyed university students from three European countries and examined their trust in two policy plans of the European Commission, and it's shaping by the appearance of its logo in the communication. They found that including the Commission’s logo significantly increased citizens' trust in these policies. Their results show an impressive effect size equivalent to 0.7 standard deviations, on average.

@@

Even though these studies substantiate a link between the appearance of symbolic elements in communications and citizens' greater trust in public organizations and policies, they did not directly investigate the causal mechanisms that underlie this effect. Therefore, we still lack understanding of the circumstances under which symbols are likely to be more/less effective in shaping citizens' trust. Specifically, these studies neither theorized nor empirically tested the interaction between symbolic elements and citizens’ processing of substantive information in communications. We do not know whether well-designed symbols reinforce, or offset, citizens' rational responses to information. Specifically, as explained above, it is essential to assess the possibility that symbols can unduly compensate for unfavorable or logically unpersuasive information that should yield negative responses.

@@

The social psychology Elaboration Likelihood Model (ELM) provides a comprehensive theoretical framework for investigating peoples' varying responses to "persuasive" communications [@petty_1986; @petty_2011]. It theorizes when and how individuals are more/less likely to change their attitudes following a communication or a message. Accordingly, ELM can be applied to various settings, types of attitudes and forms of communications. It has been specifically applied to studies on persuasion or attitudinal change in various areas, including commercial advertising [e.g. @petty_1983; @dotson_2000; @wheeler_2005; @moore_1987], public health campaigns [@wilson_2007; @angst_2009; @ruiter_2001; @briol_2006], media coverage [@petty_2002; @vanderwurff_2016] and media entertainment [@young_2008; @slater_2002]. So far however the relevance of ELM has not been directly examined in relation to changes in citizens' trust in public organizations in response to their public communications.

According to ELM, the responses of individuals to persuasive communications depends on how much effort they devote to thinking about the content of the message (or "the amount of thinking" according to ELM terminology). Persuasion, or attitudinal change, can occur when thinking is "high" or "low", but the processes and consequences are likely to differ in each situation. This range between high and low degrees of thought is referred to as the "elaboration continuum". People's location along this continuum is associated with different psychological processes of persuasion. At the high end of the elaboration continuum, also called the "central route", people are likely to respond by cognitively scrutinizing the arguments central to the merits of the issue. At the low end, also called "the peripheral route" people respond to simple cues automatically and unconsciously through primitive psychological processes that require very little cognitive effort, such as an unconscious transfer of affect[^footnote_dual_process_theories]. In turn, these two mechanisms are likely to cause divergent types of persuasion. Persuasion via the central route is expected to be more enduring, resistant and predictive of behavior, compared with the peripheral route. Of course, often people are located in between the two ends of the continuum, and thus persuasion is determined by a combination of these processes.

@@

ELM enables us to disentangle the distinct effects of different elements in communications. Some elements are processed differently by people at different locations along the elaboration continuum. Consider for instance a text that contains a series of logical arguments supporting a certain position. Those at the high end of the continuum would regard this element as argumentation. They would systematically scrutinize it and form their attitudes about the position based on its logical persuasive quality (central route processing). Stronger arguments would increase their inclination to support the position (i.e. to form positive attitudes about it). Conversely, those at the low end of the continuum are more likely to disregard the content of the text, and instead, rely on the number of lines in it, for example, as a cue for its strength (peripheral route processing). Some elements provide potential cues that lack any logical quality at all. For example, the physical attractiveness of the source normally does not represent any substantive argument, yet it can serve as a simple cue. Accordingly, they may affect the attitudes of those at the low end, but are less likely to persuade those at the high end [@petty_1986].

@@

How may we apply the theoretical framework of ELM to the study of the effects of symbolic elements in government communications on citizens' trust and their interaction with substantive information?

In light of the conceptualization of ELM, substantive information elements in government public communications represent issue-related arguments, which can be scrutinized or processed systematically based on their logical quality (via the central route), at least by those who are sufficiently motivated and able to do so. More positive information about organizations' performance and more logically persuasive arguments supporting their actions would increase citizens' trust in these organizations and their actions and vice versa. Conversely, symbolic elements are potential peripheral cues. they can increase citizens' trust via the peripheral route, by evoking positive emotions and associations that are attributed to government organizations and their policies [@alonbarkat_2017]. However, in addition to their direct effect as peripheral cues, symbolic elements may also increase citizens' trust indirectly. They can decrease the amount of thought citizens dedicate to the communication, and thus prevent them from scrutinizing substantive information that would otherwise decrease their trust, namely negative information (e.g. information on poor performance) and weak arguments (e.g. logically unpersuasive arguments supporting their actions and policies). The latter assertion entails an interaction between symbols and the substantive information in communications: symbolic elements in public communications are more likely to affect citizens' trust when communications include negative information or weak logical arguments.[^footnote_interaction_strong_information]

@@

How can symbolic elements decrease citizens' ability and motivation to think about the message? According to social psychology theory, they can do so in two main ways. First, they can cause elaboration disruption, by distracting citizens and capturing their attention and cognitive efforts [@petty_1976; @petty_1986]. Second, they can decrease citizens' motivation and ability to think about the message content by creating a positive affect and placing citizens in a positive mood. As explained above, the postulation that positive mood decreases elaboration has been explained, in part, by a mood maintenance mechanism [@mackie_1989; @bless_1990; @schwarz_1991]. This mechanism is particularly relevant to the context of government communications, in which thinking about the content of the message – government organizations and policies – is likely to yield a negative mood among most people [@marvel_2015; @marvel_2015a; @hvidman_2016]. Therefore, when citizens are exposed to symbolic elements that evoke positive feelings, they may seek to avoid thinking about tedious bureaucratic affairs, which would depress their positive mood.

@@

Keeping these limitations in mind, this study has several important implications for public administration research. Most importantly, my findings highlight the significant role of branding and symbolic communication in shaping citizens' attitudes about government. Citizens' responses to symbolic elements in communications thus far received relatively little attention in Public administration literature, compared with the abundant research on their responses to information. The latter body of literature emphasizes the cognitive biases that distort citizens' processing of information in government communications, and suggests that citizens tend to be sceptic about their positive information [@baekgaard_2016; @james_2014; @james_2017; @marvel_2015; @olsen_2017; @james_2017b; @christensen_2018; @james_2017a; @james_2017c]. Hence, it undermines the effectiveness of the communication of substantive information. Yet, this study, and others, demonstrated that symbolic communication, which requires little cognitive effort, can be fairly effective in increasing citizens' trust, and mitigating their negative attitudes [@alonbarkat_2017; @karens_2016; @teodoro_2018].

Furthermore, my findings suggest that symbols decrease citizens' "rational", systematic processing of substantive information. This is good news as far as we are concerned mainly about the negative biases of citizens and the media regarding public organizations [@hvidman_2016; @hvidman_2018; @marvel_2015a], resulting in citizens' *undervaluation* of government performance. On the other hand, the emotive effect of government symbols on public opinion is highly problematic. It challenges the democratic ideal that citizens form their opinions about their government and its actions on the basis of rational thinking, rooted in healthy skepticism. Congruently, it contrasts with the main objective of public communication to enhance bureaucracies’ accountability to citizens [@canel_2013; @lee_2012; @lee_2001; @lee_2008; @lee_2011], and with the basic moral “obligation of the public administrator to keep the public informed” [@lee_2000, 459].

Moreover, the effect of symbolic elements becomes especially problematic when they are used by poorly performing organizations, or organizations that lack professional expertise. In those cases, symbolic elements can mitigate justified public criticism of government failures. Theoretically, this distortion could be either the unconscious result of an honest attemt by the bureaucracy to promote good causes, or worse - a satrategically planned "propaganda". This could be used for promoting the narrow political interests of parties and individual politicians controlling bureaucracies - for instance, the Minister of Environment Protection and/or his party [For exmaples of politicians' exploitation of public communications see @sanders_2013a]. At the same time, it could also be used for promoting the interests of the bureaucracies themselves, namely nurturing their long-term positive reputation, which provides them with political power and autonomy [@busuioc_2016; @bertelli_2018]. In any case, it may result in undermining the accountability of bureaucrats and politicians to the public, and a deviation from the model of democratic responsiveness [@dahl_1973; @soroka_2010].

@@

Hence, whereas government communication with citizens is normally perceived as a tool for enhancing democratic responsiveness and accountability, it can also be use it as a means to undermine them. ##work on that##
